Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots

### 2022-12-06 mdpi2022roscia.tex
light blue area

% flywheel
Flywheels represent an additional option for controlling the robot orientation.
Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient the robot. This device has been first applied for spacecraft orientation \cite{oland2009reaction}, and later it has been only sporadically investigated in legged locomotion for controlling pitch orientation, both for bipeds \cite{Brown2016}, \cite{xiong2020sequential} and quadrupeds \cite{kolvenbach2019towards}, \cite{vasilopoulos2016quadruped}.
Compared with tails, flywheels do not have position limits, and since their rotation axes pass through the \gls{com}, its angular momentum results holonomic \cite{machairas2015quadruped}.
To get a fast response, it is necessary to have an abrupt change in the flywheels angular velocity (angular acceleration).
Using a brake avoids the employment of a motor able to deliver higher torques \cite{gajamohan2012cubli}, 
keeping the system compact: the motor slowly accelerates the wheel to a certain speed to store angular momentum and, when a reorientation is required, the break stops its spin.
Since the effect of the break is unidirectional, a limitation of this approach is given by the fact that it is possible to generate a rotation of the base only in the opposite direction of the flywheel angular velocity making them unsuitable for applications where \textit{continuous} controllability is required. On the other hand, direct-drive controlled flywheels enable to create accelerations in both directions,  and  to implement continuous control laws. \\
The benefits of this kind of \gls{ocs} are multiple. 
\begin{enumerate}
\item Flywheels can correct orientation errors during the flight (e.g. due to disturbances or tracking inaccuracies in the angular momentum achieved at the lift-off)   % and non-idealities).  
in a continuous manner. 
\item They enable to track a time-varying reference (e.g. to have the robot land with a desired angular velocity (possibly zero) and orientation).
\item  Even in presence of contacts, they can enhance the landing phase or stabilize dynamic gaits (e.g. trot), by significantly reducing trunk oscillations.  
\item  The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
\end{enumerate}
In more complex scenarios, like for a somersault, limbs and \gls{ocs} can operate in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=.7\linewidth]{figures/fw_jump.eps}
	\caption{\small First Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs. time plots. A disturbance moment $\bm{\tau}_{dist}$ on the trunk can be compensated only if the \gls{ocs} is enabled. If it is disabled, the robot is unable to restore a safe configuration after the touch-down and eventually falls.}
	\label{fig:fw_jump}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[width=.7\linewidth]{figures/fall_plot.eps}
	\caption{\small Second Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs. time plots. The \gls{ocs} drives the robot orientation during a fall to be horizontal. This allows to damp for damping the base oscillations after the touch-down (vertical dashed line), even without implementing a landing strategy}
	\label{fig:fall_plot}
\end{figure}

>>>>

% flywheel
Flywheels represent an additional option for controlling the robot orientation.
Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient the robot. This device has been first applied for spacecraft orientation \cite{oland2009reaction}, and later it has been only sporadically investigated in legged locomotion for controlling pitch orientation, both for bipeds \cite{Brown2016}, \cite{xiong2020sequential} and quadrupeds \cite{kolvenbach2019towards}, \cite{vasilopoulos2016quadruped}.
Compared with tails, flywheels do not have position limits, and since their rotation axes pass through the \gls{com}, its angular momentum results holonomic \cite{machairas2015quadruped}.
To get a fast response, it is necessary to have an abrupt change in the flywheels angular velocity (angular acceleration).
Using a brake avoids the employment of a motor able to deliver higher torques \cite{gajamohan2012cubli}, 
keeping the system compact: the motor slowly accelerates the wheel to a certain speed to store angular momentum and, when a reorientation is required, the break stops its spin.
Since the effect of the break is unidirectional, a limitation of this approach is given by the fact that it is possible to generate a rotation of the base only in the opposite direction of the flywheel angular velocity making them unsuitable for applications where \textit{continuous} controllability is required. On the other hand, direct-drive controlled flywheels enable to create accelerations in both directions,  and  to implement continuous control laws. \\
The benefits of this kind of \gls{ocs} are multiple. 
\begin{enumerate}
\item Flywheels can correct orientation errors during the flight (e.g. due to disturbances or tracking inaccuracies in the angular momentum achieved at the lift-off)   % and non-idealities).  
in a continuous manner. 
\item They enable to track a time-varying reference, e.g. to have the robot land with a desired angular velocity (possibly zero) and orientation.
\item  Even in presence of contacts, they can enhance the landing phase or stabilize dynamic gaits (e.g. trot), by significantly reducing trunk oscillations.  
\item  The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
\end{enumerate}
In more complex scenarios, like for a somersault, limbs and \gls{ocs} can operate in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=.7\linewidth]{figures/drawing.png}
	\caption{\small First Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs. time plots. A disturbance moment $\bm{\tau}_{dist}$ on the trunk can be compensated only if the \gls{ocs} is enabled. If it is disabled, the robot is unable to restore a safe configuration after the touch-down and eventually falls. The light blue area represents the interval of time in which the disturbance is applied.}
	\label{fig:fw_jump}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[width=.7\linewidth]{figures/fall_plot.eps}
	\caption{\small Second Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs. time plots. The \gls{ocs} drives the robot orientation during a fall to be horizontal. This allows to damp for damping the base oscillations after the touch-down (vertical dashed line), even without implementing a landing strategy}
	\label{fig:fall_plot}
\end{figure}
